# Generating mock communities {-}

Using the article by Morney et al. we identified 16 species we believed to be good candidate to include in the mock communities. Four types of mock communities were designed : 

16 species - normalized abundance (richness + / evenness +)
16 species - non-normalized  (richness + / evenness -) 
8 species - normalized abundance (richness - / evenness +) 
8 species - normalized abundance (richness - / evenness -) 

To generate the low richness communities, 8 species were randomly selected from the 16 species. 

Required libraries 
```{r, eval=FALSE}
library(dplyr)
library(stringr)
library(phyloseq)
```

Generate and load required files (list of fishes and databases sequence headers)
```{r, eval=FALSE}
rich_plus = c("Salvelinus_fontinalis","Oncorhynchus_mykiss","Lota_lota","Culaea_inconstans","Thymallus_arcticus","Perca_flavescens","Coregonus_clupeaformis","Prosopium_williamsoni","Salmo_trutta","Sander_vitreus","Couesius_plumbeus","Platygobio_gracilis","Coregonus_artedi","Catostomus_catostomus","Percopsis_omiscomaycus","Cottus_cognatus")
# Low richness community 
# Random subsample to ID which fishes to include in the low richness community 
rich_low = sample(x=fish_id_bad, size = 8)
rich_low
#"Salvelinus_fontinalis","Oncorhynchus_mykiss","Lota_lota","Culaea_inconstans","Thymallus_arcticus","Perca_flavescens","Coregonus_clupeaformis","Prosopium_williamsoni"
```

:::::::::::::::::::::::::::::::: 12S marker gene ::::::::::::::::::::::::::::::::

Extract sequence header name 
```{bash, eval = FALSE}
grep -e ">" /home/kvilleneuve/fish_edna/database/curated12s_db/12s_alberta_curated_combined.fasta > /home/kvilleneuve/fish_edna/database/curated12s_db/12s_curated_header.txt
```

Generate read count file for each type of communities

Subset sequence header to keep only fish from generated list
```{r, eval=FALSE}
# Load sequence header from database 
header12s = read.table("/home/kvilleneuve/fish_edna/database/curated12s_db/12s_alberta_curated_combined_ready.fasta")
# Set reads per samples (number obtain from the DADA2 workflow)
average_reads = 80492

############################################################
##########  High Richness and high evenness community ###########
############################################################

# Subset sequence header to keep only fish from generate list 
header12s_richHIGH = header12s %>% 
  filter(if_any(1, str_detect, paste0(rich_plus, collapse = '|')))
# Remove > at the beginning 
header12s_richHIGH$V1 = gsub(">", "", header12s_richHIGH$V1)
# Save list 
write.table(header12s_richHIGH, "/home/kvilleneuve/fish_edna/raw/mock_comm/12s_header_richplus.txt", quote = FALSE, col.names = FALSE, row.names = FALSE)

# Generate read count file for (richness + / evenness +) mock community. We want the sum of all the ASV in our sample to be similar to the average read count per sample in the regular samples. 
header12s_richHIGH$count = floor(average_reads/nrow(header12s_richHIGH))
write.table(header12s_richHIGH, "/home/kvilleneuve/fish_edna/raw/mock_comm/12s_counts_richplusevenplus.txt", quote = FALSE, col.names = FALSE, row.names = FALSE, sep = "\t")

###########################################################
##########  Low Richness and high evenness community ###########
###########################################################

header12s_richLOW = header12s %>% 
  filter(if_any(1, str_detect, paste0(rich_low, collapse = '|')))
# Remove > at the beginning 
header12s_richLOW$V1 = gsub(">", "", header12s_richLOW$V1)
# Save list 
write.table(header12s_richLOW, "/home/kvilleneuve/fish_edna/raw/mock_comm/12s_header_richlow.txt", quote = FALSE, col.names = FALSE, row.names = FALSE)

# Generate readcount_file for  (richness - / evenness +) mock community 
header12s_richLOW$count = floor(average_reads/nrow(header12s_richLOW))

write.table(header12s_richLOW, "/home/kvilleneuve/fish_edna/raw/mock_comm/12s_counts_richlowevenplus.txt", quote = FALSE, col.names = FALSE, row.names = FALSE, sep = "\t")

###########################################################
###############  Low evenness communities #################
###########################################################

# The low-evenness of both richness types of mock community was manually 
# generated to have an exponential distribution of reads alternating between 
# non- and consistently identified fishes.

# 12s_counts_richplusevenlow.txt
# 12s_counts_richlowevenlow.txt
```

**Create FASTA for InSilicoSeq**

Generate FASTA file with sequences from all 16 fishes used to generate mock community.

Because InSilicoSeq only keeps the first 300 nucleotides, the generated amplicon sequence did not initially correspond to our amplified region and no merged reads were generated. I are therefore used cutadapt to remove all the nucleotides before the forward primer while retaining the primer. The same methodology was used for the reverse primer in order to generate a reverse sequence. We then need to seperately generate the forward and reverse reads

```{bash, eval=FALSE}
# Forward primer 
cutadapt --action=retain -g CCGGTAAAACTCGTGCCAGC -o 12s_seqTRIMF.fasta 12S_allnormline.fasta 
# Reverse primer
cutadapt --action=retain -a CAAACTGGGATTAGATACCCCACTATG -o 12s_seqTRIMR.fasta  12S_allnormline.fasta
```

**Use InSilicoSeq**

Use [InSilicoSeq](https://insilicoseq.readthedocs.io/en/latest/index.html) pre-computed error models to generate amplicon reads. Illumina instruments : 

```{bash, eval=FALSE}
#!bin/bash
############# High Richness and High Evenness ############# 
# FORWARD read
echo "Generating reads for High Richness and High Evenness"
iss generate --genomes 12s_seqTRIMF.fasta --readcount_file  12s_counts_richplusevenplus.txt --sequence_type amplicon --model MiSeq  --output 12s-richplusevenplus_L001 --cpus 20
# Delete the reverse (R2) and move R1 to folder fastq
rm *R2* 
mv *.fastq fastq 
# REVERSE  read
iss generate --genomes 12s_seqTRIMR.fasta --readcount_file  12s_counts_richplusevenplus.txt --sequence_type amplicon --model MiSeq  --output 12s-richplusevenplus_L001 --cpus 20
rm *R1*
mv *.fastq fastq 

############# Low Richness and high evenness ############# 
echo "Generating reads for Low Richness and High Evenness"
# FORWARD read
iss generate --genomes 12s_seqTRIMF.fasta --readcount_file  12s_counts_richlowevenplus.txt --sequence_type amplicon --model MiSeq  --output 12s-richlowevenplus_L001 --cpus 20
# Delete the reverse (R2) and move R1 to folder fastq
rm *R2* 
mv *.fastq fastq 
# REVERSE  read
iss generate --genomes 12s_seqTRIMR.fasta --readcount_file  12s_counts_richlowevenplus.txt --sequence_type amplicon --model MiSeq  --output 12s-richlowevenplus_L001 --cpus 20
# Delete  the forward (R1) and move R2 to folder fastq
rm *R1*
mv *.fastq fastq 

############# High Richness and low evenness ############# 
echo "Generating reads for High Richness and Low Evenness"
# FORWARD read
iss generate --genomes 12s_seqTRIMF.fasta --readcount_file 12s_counts_richplusevenlow.txt --sequence_type amplicon --model MiSeq  --output 12s-richplusevenlow_L001 --cpus 20
# Delete all the reverse (R2) and move R1 to another folder (fastq)
rm *R2* 
mv *.fastq fastq 
# REVERSE  read
iss generate --genomes 12s_seqTRIMR.fasta --readcount_file  12s_counts_richplusevenlow.txt --sequence_type amplicon --model MiSeq  --output 12s-richplusevenlow_L001 --cpus 20
# Delete all the forward (R1) and move R2 to another folder (fastq)
rm *R1*
mv *.fastq fastq 

############# Low Richness and low evenness ############# 
echo "Generating reads for Low Richness and Low Evenness"
# FORWARD read
iss generate --genomes 12s_seqTRIMF.fasta --readcount_file 12s_counts_richlowevenlow.txt --sequence_type amplicon --model MiSeq  --output 12s-richlowevenlow_L001 --cpus 20
# Delete all the reverse (R2) and move R1 to another folder (fastq)
rm *R2* 
mv *.fastq fastq 
# REVERSE  read
iss generate --genomes 12s_seqTRIMR.fasta --readcount_file  12s_counts_richlowevenlow.txt --sequence_type amplicon --model MiSeq  --output 12s-richlowevenlow_L001 --cpus 20
# Delete all the forward (R1) and move R2 to another folder (fastq)
rm *R1*
mv *.fastq fastq 

rm *.vcf
```

The pattern `_001` was then manually added to every fastq generated to match pattern from our other samples. 

:::::::::::::::::::::::::::::::: COI marker gene ::::::::::::::::::::::::::::::::

Extract sequence header name 
```{bash, eval = FALSE}
grep -e ">" /home/kvilleneuve/fish_edna/database/curatedcoi_db/coi_alberta_curated_NCBI.fasta > /home/kvilleneuve/fish_edna/database/curatedcoi_db/coi_NCBI_header.txt
```

Generate read count files for each type of communities 
```{r, eval = FALSE}
#  Load Sequence header from databases
headercoi = read.table("/home/kvilleneuve/fish_edna/database/curatedcoi_db/coi_NCBI_header.txt")
# Set reads per samples (number obtain from the DADA2 workflow)
average_reads = 55424

############################################################
##########  High Richness and evenness community ###########
############################################################

# High richness community 
rich_plus = c(fish_id_good, fish_id_bad)

# Subset sequence header to keep only fish from generated list
headercoi_richHIGH = headercoi %>% 
  filter(if_any(1, str_detect, paste0(rich_plus, collapse = '|')))

headercoi_richHIGH$V1 = gsub(">", "", headercoi_richHIGH$V1)

write.table(headercoi_richHIGH, "/home/kvilleneuve/fish_edna/raw/mock_comm/coi_header_richplus.txt", quote = FALSE, col.names = FALSE, row.names = FALSE, sep = ";")

# Generate read count file for (richness + / evenness +) mock community. We want the sum of all the ASV in our sample to be similar to the average read count per sample in the regular samples. 
headercoi_richHIGH$count = floor(average_reads/nrow(headercoi_richHIGH))
write.table(headercoi_richHIGH, "/home/kvilleneuve/fish_edna/raw/mock_comm/coi_counts_richplusevenplus.txt", quote = FALSE, col.names = FALSE, row.names = FALSE, sep = "\t")

###########################################################
##########  Low Richness and evenness community ###########
###########################################################

# Low richness community 
rich_low = c("Oncorhynchus_mykiss_x_Salmo_salar","Notropis_atherinoides","Percina_caprodes","Coregonus_artedi","Sander_vitreus","Percopsis_omiscomaycus","Culaea_inconstans","Lota_lota")

# Subset sequence header to keep only fish from generated list
headercoi_richLOW = headercoi %>% filter(if_any(1, str_detect, paste0(rich_low, collapse = '|')))

headercoi_richLOW$V1 = gsub(">", "", headercoi_richLOW$V1)

write.table(headercoi_richLOW, "/home/kvilleneuve/fish_edna/raw/mock_comm/coi_header_richlow.txt", quote = FALSE, col.names = FALSE, row.names = FALSE)

# Generate read count file for (richness - / evenness +) mock community.  

headercoi_richLOW$count = floor(average_reads/nrow(headercoi_richLOW))

write.table(headercoi_richLOW, "/home/kvilleneuve/fish_edna/raw/mock_comm/coi_counts_richlowevenplus.txt", quote = FALSE, col.names = FALSE, row.names = FALSE, sep = "\t")

###########################################################
###############  Low evenness communities #################
###########################################################

# The low-evenness of both richness types of mock community was manually 
# generated to have an exponential distribution of reads alternating between 
# non- and consistently identified fishes.
```

**Create FASTA for InSilicoSeq**

Generate FASTA file with sequences from all 16 fishes used to generate mock community and make this fasta linear 

```{bash, eval = FALSE}
seqkit grep -nrf coi_header_richplus.txt
```

Because InSilicoSeq only keeps the first 300 nucleotides, the generated amplicon sequence did not initially correspond to our amplified region and no merged reads were generated. I therefore used cutadapt to remove all the nucleotides before the forward primer while retaining the primer. 

For the reverse primer, the actual primer sequence was not consistently present in our sequences. Increasing the -e value did allow the primer to be found but this resulted in sequences less then 300 bp for some fishes. I therefore decided to instead cut the trimF fasta at 300 bp. 

```{bash, eval = FALSE}
cutadapt --action=retain -g GTATTTGGYGCYTGRGCCGGRATAGT -o coi_seq_TRIMF.fasta coi_all.fasta -e 0.2
cutadapt --action=retain -a CGAATRAAYAAYATRAGYTTYTG -o coi_seq_TRIMR.fasta coi_all.fasta -e 0.2

#cutadapt --length 350 -o coi_seq_TRIMFR.fasta  coi_seq_TRIMF.fasta
```

**InSilicoSeq**

Use [InSilicoSeq](https://insilicoseq.readthedocs.io/en/latest/index.html) pre-computed error models to generate amplicon reads. 

```{bash, eval = FALSE}
# High Richness and high evenness
iss generate --genomes coi_seq_TRIMFR.fasta --readcount_file  coi_counts_richplusevenplus.txt --sequence_type amplicon --model MiSeq  --output coi-richplusevenplus_L001

# High Richness and low evenness
iss generate --genomes coi_seq_TRIMFR.fasta --readcount_file coi_counts_richplusevenlow.txt --sequence_type amplicon --model MiSeq  --output coi-richplusevenlow_L001

# Low Richness and high evenness
iss generate --genomes coi_seq_TRIMFR.fasta --readcount_file  coi_counts_richlowevenplus.txt --sequence_type amplicon --model MiSeq  --output coi-richlowevenplus_L001

# Low Richness and low evenness
iss generate --genomes coi_seq_TRIMFR.fasta --readcount_file coi_counts_richlowevenlow.txt --sequence_type amplicon --model MiSeq  --output coi-richlowevenlow_L001
```

The pattern `_001` was then manually added to every FASTQ generated to match pattern from our other samples. 

```{bash, eval = FALSE}
#!/bin/bash
############# High Richness and High Evenness ############# 
# FORWARD read
echo "Generating reads for High Richness and High Evenness"
iss generate --genomes coi_seq_TRIMF.fasta --readcount_file  coi_counts_richplusevenplus.txt --sequence_type amplicon --model MiSeq  --output coi-richplusevenplus_L001 --cpus 20
# Delete the reverse (R2) and move R1 to folder fastq
rm *R2* 
mv *.fastq fastq 
# REVERSE  read
iss generate --genomes coi_seq_TRIMR.fasta --readcount_file  coi_counts_richplusevenplus.txt --sequence_type amplicon --model MiSeq  --output coi-richplusevenplus_L001 --cpus 20
rm *R1*
mv *.fastq fastq 

############# Low Richness and high evenness ############# 
echo "Generating reads for Low Richness and High Evenness"
# FORWARD read
iss generate --genomes coi_seq_TRIMF.fasta --readcount_file  coi_counts_richlowevenplus.txt --sequence_type amplicon --model MiSeq  --output coi-richlowevenplus_L001 --cpus 20
# Delete the reverse (R2) and move R1 to folder fastq
rm *R2* 
mv *.fastq fastq 
# REVERSE  read
iss generate --genomes coi_seq_TRIMR.fasta --readcount_file  coi_counts_richlowevenplus.txt --sequence_type amplicon --model MiSeq  --output coi-richlowevenplus_L001 --cpus 20
# Delete  the forward (R1) and move R2 to folder fastq
rm *R1*
mv *.fastq fastq 

############# High Richness and low evenness ############# 
echo "Generating reads for High Richness and Low Evenness"
# FORWARD read
iss generate --genomes coi_seq_TRIMF.fasta --readcount_file coi_counts_richplusevenlow.txt --sequence_type amplicon --model MiSeq  --output coi-richplusevenlow_L001 --cpus 20
# Delete all the reverse (R2) and move R1 to another folder (fastq)
rm *R2* 
mv *.fastq fastq 
# REVERSE  read
iss generate --genomes coi_seq_TRIMR.fasta --readcount_file  coi_counts_richplusevenlow.txt --sequence_type amplicon --model MiSeq  --output coi-richplusevenlow_L001 --cpus 20
# Delete all the forward (R1) and move R2 to another folder (fastq)
rm *R1*
mv *.fastq fastq 

############# Low Richness and low evenness ############# 
echo "Generating reads for Low Richness and Low Evenness"
# FORWARD read
iss generate --genomes coi_seq_TRIMF.fasta --readcount_file coi_counts_richlowevenlow.txt --sequence_type amplicon --model MiSeq  --output coi-richlowevenlow_L001 --cpus 20
# Delete all the reverse (R2) and move R1 to another folder (fastq)
rm *R2* 
mv *.fastq fastq 
# REVERSE  read
iss generate --genomes coi_seq_TRIMR.fasta --readcount_file  coi_counts_richlowevenlow.txt --sequence_type amplicon --model MiSeq  --output coi-richlowevenlow_L001 --cpus 20
# Delete all the forward (R1) and move R2 to another folder (fastq)
rm *R1*
mv *.fastq fastq 

rm *.vcf
```
for f in *.fastq; do mv "$f" "$(echo "$f" | sed s/IMG/VACATION/)"; done
